\documentclass[fontsize=11pt, paper=a4]{scrartcl}
\usepackage[twoside=false, texlive=2020]{ubonn-thesis}
\usepackage{thesis_defs}
\usepackage{atlasmisc}
\usepackage{atlasparticle}
\usepackage{atlasunit}
\usepackage[margin=2.5cm]{geometry}

\title{Search for Higgs Boson Pair Production in the\\$\bbbar\tautau$ Final State
  with the ATLAS\\Experiment at the LHC}
\subtitle{Dissertation Summary}
\author{Christopher Deutsch}
\date{}

\begin{document}
\maketitle
\thispagestyle{empty}

\noindent A search for Higgs boson pair (\HH) production in final states with
two $b$~quarks and two \tauleptons using \SI{139}{\per\femto\barn} of
proton--proton collision data at a centre-of-mass energy of \SI{13}{\TeV}
recorded by the ATLAS detector at the Large Hadron Collider (LHC) is
presented. Searches for \HH production are vital for solidifying our
understanding of electroweak symmetry breaking in the Standard Model (SM) since
they serve as direct probes of the Higgs boson self-coupling. In addition, they
provide sensitivity to new phenomena introduced by theories beyond the SM.

The search presented in this thesis targets channels with one or more hadronic
\tauleptonC decays. It relies on effective \tauid algorithms, which are
algorithms that distinguish signatures of hadronic \tauleptonC decays in the
ATLAS detector from other sources, such as quark- or gluon-initiated jets. A
novel \tauid algorithm based on recurrent neural networks is introduced. It
exploits information from reconstructed charged-particle tracks, topological
clusters of calorimeter cell signals, and purposefully constructed
discriminating variables for \tauid. The new technique outperforms the method
previously employed in the ATLAS experiment.

Three different types of Higgs boson pair production are considered in this
thesis:~non-resonant \HH production predicted by the SM (SM~\HH production);
resonant \HH production via scalar, narrow-width resonances; and non-resonant
\HH production with anomalous values of the Higgs boson self-coupling
constant~\lambdahhh.

The search for SM~\HH production targets the gluon--gluon fusion and vector
boson fusion production modes. No statistically significant deviation from the
background-only hypothesis is observed; therefore, upper limits are set on the
SM~\HH production cross section. The search yields an upper limit of $4.7$ times
the SM cross section prediction at \SI{95}{\percent} confidence level (CL).

The search for resonant \HH production considers intermediate resonances with
masses ranging from \SIrange{251}{1600}{\GeV}. Over the considered mass range,
the largest deviation from the background-only hypothesis is observed for a mass
of \SI{1000}{\GeV} with a local (global) significance of $3.1\sigma$
($2.0\sigma$). Upper limits at \SI{95}{\percent}~CL are set on the resonant \HH
production cross section, which range from \SIrange{20}{900}{\femto\barn}
depending on the mass of the resonance.

Finally, the search for SM~\HH production is reinterpreted in the context of
anomalous values of the Higgs boson self-coupling constant. Upper limits at
\SI{95}{\percent}~CL are set on the non-resonant \HH production cross section as
a function of the self-coupling modifier
$\klambda = \lambdahhh / \lambdahhh^{\text{SM}}$, where $\lambdahhh^{\text{SM}}$
is the Higgs boson self-coupling predicted by the SM. The upper limits on the
non-resonant \HH production cross section are compared to theory predictions,
providing constraints on the self-coupling modifier of
$\num{-2.4} \leq \klambda \leq 9.2$.

\end{document}
