% ------------------------------------------------------------------------------
\chapter*{Abstract}
% ------------------------------------------------------------------------------
% Aim for 500-600 words

A search for Higgs boson pair (\HH) production in final states with two
$b$-quarks and two \tauleptons using \SI{139}{\per\femto\barn} of proton--proton
collision data at a centre-of-mass energy of \SI{13}{\TeV} recorded by the ATLAS
detector at the Large Hadron Collider
% in 2015--2018
is presented. Searches for Higgs boson pair production are vital for solidifying
our understanding of electroweak symmetry breaking in the Standard Model (SM)
since they serve as direct probes of the Higgs boson self-coupling. In addition,
they can provide sensitivity to new physical phenomena introduced by theories
beyond the SM (BSM).
% ., further motivating searches for \HH production.

The search presented in this thesis targets channels with one or more hadronic
\taulepton decay. As a result, the search relies on effective \tauid algorithms,
which are algorithms that distinguish signatures of hadronic \taulepton decays
in the ATLAS detector from those of other sources, such as quark- or
gluon-initiated jets. For this purpose, a novel \tauid algorithm based on
recurrent neural networks is introduced. The algorithm exploits information of
reconstructed charged-particle tracks, topological clusters of calorimeter cell
signals, and purposefully constructed discriminating variables for \tauid. The
new technique significantly outperforms the method previously employed in the
ATLAS experiment and has since become the default algorithm for \tauid.

In this thesis, three different types of Higgs boson pair production are
considered: non-resonant \HH production predicted by the SM (SM~\HH production);
resonant \HH production via scalar, narrow-width resonances; and non-resonant
\HH production with anomalous values of the Higgs boson self-coupling
constant~\lambdahhh. In all cases, machine learning methods are used to separate
signal from background events.

The search for SM~\HH production targets the gluon--gluon fusion (\ggF) and
vector boson fusion (VBF) production modes. No statistically significant
deviation from the background-only hypothesis is observed; therefore, upper
limits are set on the SM~\HH production cross section \xsecggfvbf and signal
strength~$\mu = \xsecggfvbf / \xsecggfvbf^{\text{SM}}$, where
$\xsecggfvbf^{\text{SM}}$ is the SM cross section prediction. The search yields
upper limits of $\mu < \num{4.7}$ and $\xsecggfvbf < \SI{140}{\femto\barn}$ at
\SI{95}{\percent} confidence level.

Resonant \HH production via new scalar resonances~$X$ is predicted by several
BSM theories, for example, theories involving an extended Higgs sector. A search
for scalar, narrow-width resonances with masses ranging from
\SIrange{251}{1600}{\GeV} is performed. Over the considered mass range, the
largest excess is observed for a resonance mass of \SI{1000}{\GeV} with a local
(global) significance of $3.1\sigma$ ($2.0\sigma$). Upper limits are set on the
$\pp \to X \to \HH$ production cross section, which range from
\SIrange{20}{900}{\femto\barn} depending on the mass of the resonance.

Finally, the search for SM~\HH production is reinterpreted in terms of anomalous
values of the Higgs boson self-coupling constant. Upper limits at
\SI{95}{\percent} confidence level are set on the non-resonant \HH production
cross section as a function of the self-coupling modifier \klambda, which is
defined as $\klambda = \lambdahhh / \lambdahhh^{\text{SM}}$, where
$\lambdahhh^{\text{SM}}$ is the self-coupling predicted by the SM. The upper
limits on the non-resonant \HH production cross section is compared to the
theoretical predictions yielding an allowed \klambda interval of
$\num{-2.4} \leq \klambda \leq 9.2$.

% Examples: Ray
% https://bonndoc.ulb.uni-bonn.de/xmlui/handle/20.500.11811/9447
%
% Tobias
% https://bonndoc.ulb.uni-bonn.de/xmlui/handle/20.500.11811/9266


%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../phd_thesis"
%%% End:
