% ----------------------------------------------------------------------------
\chapter*{Acknowledgements}
% ----------------------------------------------------------------------------

The completion of this dissertation would not have been possible without the
support and encouragement of many people. I am deeply grateful for the
contributions of those who have been a part of my journey. Although it is not
possible to mention each person by name, I would like to extend my gratitude to
all those who have helped to make this thesis a reality.

First of all, I am deeply indebted to Prof.\ Dr.\ Jochen Dingfelder for
accepting me as part of his research group and giving me the freedom to pursue
many interesting topics during my PhD. I am grateful for your unwavering
guidance, support, and patience throughout my MSc and PhD studies in your
group. \emph{Thank you!}

I am extremely grateful to my co-supervisor Tatjana Lenz. Thank you for
supporting me and for the many occasions where you proofread the things I
wrote, be it abstracts, internal notes, or chapters of this thesis. Thank you
also for coordinating the $\HH \to \bbtautau$ effort in ATLAS, which, I am sure,
must have felt like herding cats at times.

Many thanks to Prof.\ Dr.\ Ian Brock for accepting to be a referee for my
dissertation. Moreover, I am grateful to PD Dr.\ Bastian Kubis and Prof.\ Dr.\
Elena Demidova for agreeing to be members of my doctoral committee.

My appreciation goes out to the members and alumni of the \textsc{HEP1}
group. Thank you, Shubham Bansal, Florian Beisiegel, Christina Dimitriadi,
Stephan Duell, Peter Falke, Stephan Hageböck, Christian Nass, Oğul Öncel,
Jonathan Pampel, Christos Vergis, and Eckhard von Törne for the lunches, coffee
breaks, and fruitful discussion. It was a pleasure to be able to work with all
of you. In addition, I am thankful to Shubham, Christina, Peter, Christian, and
Eckhard for giving me feedback on parts of this thesis.

Special thanks go to Benedict Winter, William Davey, Michel Janus, and
Pier-Olivier DeViveiros for taking me under their wings during the early stages
of my career in the ATLAS collaboration. I appreciate your valuable advice and
infectious enthusiasm, which was a big part in motivating me to continue on the
academic track after finishing my Master's thesis.

I would like to acknowledge my collaborators at the ATLAS experiment: Alessandra
Betti, Christian Grefe, Zhiyuan ``Jordan'' Li, Yanlin Liu, Max Märker, Bertrand
Martin Dit Latour, Tulin Mete, Giacinto Piacquadio, Elliot Reynolds, Soshi
Tsuno, Song-Ming Wang, Daniele Zanzi, and Bowen Zhang. Many thanks also to my
\bbtautau friends Petar Bokan, Christina Dimitriadi, and Serhat Ördek for the
post-work Zoom hangouts during the pandemic and introducing me to the art of
crab walking.

Lastly, I want to thank my parents. None of this would have been possible
without your unparalleled support. Thanks should also go to Hilda for
occasionally taking my mind off the struggles of doing a PhD.

\vspace{2\baselineskip}

\hfill%
\begin{tabular}{l}
  Christopher Deutsch \\
  Bonn, February 2023
\end{tabular}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../phd_thesis"
%%% End: 
