The estimation of \jettotauhadvis backgrounds in the \lephad channels is
outlined in the following. A data-driven background estimation method yielding a
combined estimate of the multi-jet and \ttbar background with a \faketauhadvis
is adopted. The method is an extension of the fake factor method, previously
introduced in \Cref{sec:hadhad_multijet}, to account for multiple sources of
\faketauhadvis that can differ in their process-specific fake factors.\todo{How
  does \Wjets fit in here?}

Events in the \lephad channel where the \tauhadvis candidate is an
\antitau define the Anti-ID region used for the fake factor
method. Two control regions are defined that are enhanced in multi-jet
and \ttbar events, respectively. Both regions have a corresponding ID
and Anti-ID region and are orthogonal to the SRs. These CRs are used
to determine fake factors specifically for \faketauhadvis from
multi-jet and \ttbar events. The CR definitions and fake factor
calculations are described in the following:
\begin{description}

\item[Multi-jet fake factors] are determined in a region defined by
  the requirement that the electron/muon fails the loose isolation
  working point and removing the $\mBB < \SI{150}{\GeV}$ selection
  applied in the SRs. The remainder of the selection is identical to
  the SR selection of the \lephad channels. This CR has high multi-jet
  purity and allows to calculate multi-jet fake factors, \FFqcd, as
  the ratio of multi-jet events in the ID and Anti-ID region. The
  number of multi-jet events is estimated by subtracting the number of
  non-multi-jet events estimated using simulation from the observed
  number of events in a given region.

\item[\ttbar fake factors] are determined in a region defined by
  requiring $\mBB > \SI{150}{\GeV}$ while other selections remain
  identical to the SR selection. This CR has high \ttbar purity but is
  not pure in \ttbarFakes events
  % relevant for the fake factor calculation
  due to contributions of \ttbar events with \truetauhadvis that have
  to be subtracted. Fake factors for \ttbar, \FFttbar, are calculated
  as the ratio of \ttbarFakes events in ID and Anti-ID regions. The
  number of \ttbarFakes events is estimated, assuming negligible
  contribution of multi-jet events, by subtracting the number of
  non-\ttbarFakes events estimated using simulation from the observed
  number of events in a given region.

\end{description}
The fake factor calculation is performed separately for the \lephad
SLT and LTT channels, and separately for 1- and 3-prong \tauhadvis
candidates. Moreover, the fake factors are measured in bins of
\tauhadvis candidate \pT.

To estimate the \faketauhadvis background in the SRs, it needs to be
considered that the Anti-ID regions corresponding to the SRs consist
of mixtures of multi-jet and \ttbar events with \faketauhadvis. Let
\rqcd be the fraction of \faketauhadvis events from multi-jet
processes in this Anti-ID region. \emph{Combined fake factors} are
defined as the weighted combination of \FFqcd and \FFttbar:
\begin{align*}
  \FFcomb = \rqcd \, \FFqcd + (1 - \rqcd) \, \FFttbar \,\text{.}
\end{align*}
The combined fake factors can be applied to events with
\faketauhadvis, irrespective of whether the event originates from
multi-jet or \ttbar processes, to yield the background estimate in the
ID region. The determination of \rqcd proceeds according to
\begin{align}
  \rqcd =
  \frac{N_{\text{multi-jet}}}{N_{\text{multi-jet}} + N_{\text{\ttbarFakes}}}
  = \frac
  { N_{\text{data}} - N_{\text{non-multi-jet}} }
  { N_{\text{data}} - N_{\text{non-(multi-jet or \ttbarFakes)}} } \,\text{,}
  \label{eq:rqcd}
\end{align}
where $N_{\text{data}}$ is the number of events observed in the
Anti-ID region and $N_{p}$ the number of events expected from process
$p$ in the Anti-ID region. The subtractions on the right-hand side of
\Cref{eq:rqcd} use the expected number of events predicted using
simulation. This includes the subtraction of \ttbarFakes events in the
numerator.

The determination of \rqcd is performed separately for the \lephad SLT
and LTT channels, 1- and 3-prong \tauhadvis candidates, and events
containing electrons and muons. In addition, \rqcd is determined in
bins of \pT of the \tauhadvis candidate. In the SLT channel, \rqcd is
typically small or zero showing that the majority of \faketauhadvis
backgrounds originate from \ttbar. The multi-jet contribution in the
LTT channel is larger with \rqcd ranging from
\SIrange{10}{30}{\percent} depending on \tauhadvis candidate \pT and
\Ntracks. Uncertainties on the \rqcd estimate have little impact on
the \faketauhadvis background prediction since \FFqcd and \FFttbar
tend to be of similar size in most bins.

The \faketauhadvis estimation in the \lephad channel is concluded by
briefly discussing the reason for not using a combined fake factor
method in the \hadhad channel. In general, using a similar method in
the \hadhad channel would be preferred compared to performing separate
estimates of the \faketauhadvis backgrounds from multi-jet and \ttbar.
This is because the combined fake factor method does not require to
distinguish between events with \faketauhadvis from multi-jet and
\ttbar when applying the fake factors to events from the Anti-ID
region. Currently, the multi-jet estimate in the \hadhad channel
requires a large subtraction of \ttbar events with \faketauhadvis
which is a large source of systematic uncertainty on the multi-jet
estimate. For the combined fake factor method, this uncertainty would
be restricted to an uncertainty on the \rqcd determination. However,
under the assumption that \FFqcd and \FFttbar are similar, a large
uncertainty on \rqcd would only have a small effect on the resulting
\faketauhadvis background estimate. Difficulties arise when trying to
adopt the method to the \hadhad channel which were previously
discussed in \Cref{sec:bkg_hadhad_ttbarfakes}. It is possible that
these can be overcome in the future. At the current stage the search
is not limited by uncertainties related to the \faketauhadvis
estimation methods and therefore this approach was not
pursued. However, the method should be considered for future
iterations of this search for which systematic uncertainties will
become more relevant.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../../phd_thesis"
%%% End:
