\section{Background Estimation}
\label{sec:background_estimation}

The dominant background processes for the search for Higgs boson pair
production in the $\bbtautau$ final state are \ttbar, \Zjets, and
backgrounds with \tauhadvis candidates from misidentified quark- or
gluon-initiated jets, hereafter referred to as
\faketauhadvis. Backgrounds with \faketauhadvis primarily orginate
from multi-jet, \Wjets, and \ttbar processes. Minor contributions
arise from single-top, $\ttbar+V$, diboson, and production of single
Higgs bosons decaying into $\tautau$ or \bbbar. The single Higgs boson
production modes considered in this search are \ggF, VBF, \VH and \ttH
for $H \to \tautau$; $\VH$ and $\ttH$ for $H \to \bbbar$. While single
Higgs boson production is classified as a minor background here, it
has relevance to the extraction of \HH signals due to their similar
signatures.

The \Zjets and \ttbar backgrounds are estimated using templates
obtained from simulation with their normalisations being determined in
a simultaneous likelihood fit to observed data in signal and control
regions. A control region enriched in $Z$ boson production in
association with jets from quarks of heavy flavour (\ZHF) is defined
in \Cref{sec:bkg_zjets} providing constraints on the normalisation of
this background. The signal region of the \lephad SLT channel serves
to constrain the normalisation of the \ttbar background as it selects
\ttbar with high purity.

Major backgrounds containing at least one \faketauhad are estimated
using (semi-)data-driven methods. In the \hadhad channel, the primary
sources of \faketauhadvis are \ttbar and multi-jet production for
which separate estimation techniques are used. The \faketauhadvis
background from \ttbar (\ttbarFakes), where events contain either one
or two \faketauhadvis, is estimated by applying scale factors to
\faketauhadvis in simulated \ttbar events that correct the
\jettotauhadvis misidentification efficiencies in simulation using
scale factors measured in a top control region. This method will be
described in \Cref{sec:bkg_hadhad_ttbarfakes}. In contrast, the
multi-jet background, where both \tauhadvis are \faketauhadvis, is
estimated using a fully data-driven \emph{Fake Factor method} that
will be introduced in \Cref{sec:bkg_hadhad_ff}.  The estimation of the
\faketauhadvis backgrounds in the \lephad channels uses a
\emph{Combined Fake Factor method} that jointly estimates the
multi-jet, \Wjets, and \ttbar backgrounds with at least one
\faketauhadvis. This method will be briefly summarised
in~\Cref{sec:bkg_lephad_combined_ff}.

The minor background contributions are estimated using simulation and
are normalised to the integrated luminosity of the collected data
using the highest order cross-section predictions available. These
backgrounds are not discussed in detail in this section, however,
theoretical uncertainties from the modelling of these processes using
simulation are discussed at a later stage in
\Cref{sec:modelling_uncertainties}.

\todo[inline]{Validation regions?}

\todo[inline]{Discuss the lack of pure validation regions particularly
  for \ttbar fakes?}

\todo[inline]{Maybe say something about \faketauhadvis backgrounds
  being rather small and very non-signal-like, thus we are not very
  sensitive to large uncertainties on these backgrounds.}

\subsection{Associated production of $\PZ \to \ell\ell$ with quarks of heavy flavour}%
\label{sec:bkg_zjets}
\input{chapters/higgs_pair_production/background_estimation_zjets.tex}

\subsection{Fake-\tauhadvis background from \ttbar production in the \hadhad channel}%
\label{sec:bkg_hadhad_ttbarfakes}
\input{chapters/higgs_pair_production/background_estimation_hadhad_ttbarfakes.tex}

\subsection{Fake-\tauhadvis background from multi-jet production in the \hadhad channel}%
\label{sec:bkg_hadhad_ff}
\input{chapters/higgs_pair_production/background_estimation_hadhad_mj.tex}

\subsection{Fake-\tauhadvis backgrounds in the \lephad channel}%
\label{sec:bkg_lephad_combined_ff}
\input{chapters/higgs_pair_production/background_estimation_lephad.tex}


%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../../phd_thesis"
%%% End:
