This chapter presents a search for Higgs boson pair production in the \bbtautau
final state using \SI{139}{\per\femto\barn} of \pp~collision data recorded at
the ATLAS experiment in Run~2 of the LHC. Non-resonant production of Higgs boson
pairs predicted by the SM (SM~\HH production) and resonant production via scalar
resonances of narrow width are considered in the search. Final states with two
hadronic \tauleptonC decays ($\bbbar\hadhad$) and with one hadronic and one
leptonic \tauleptonC decay ($\bbbar\lephad$) are targeted. These cover the
majority of decays of the \tautau system with fractions of
$\mathcal{B}(\tautau \to \hadhad) = \SI{42.0}{\percent}$ and
$\mathcal{B}(\tautau \to \lephad) =
\SI{45.6}{\percent}$~\cite{Zyla:2020zbs}. Final states with two leptonic
\tauleptonC decays, which have a branching fraction of
$\mathcal{B}(\tautau \to \taulep\taulep) =
\SI{12.4}{\percent}$~\cite{Zyla:2020zbs}, are not considered in this search.
The primary focus of this chapter lies on the \hadhad channel.

The $\bbtautau$ channel is among the most promising channels to probe the nature
of SM~\HH production. This is due to a substantial fraction of decays of Higgs
boson pairs to \bbtautau final states of about \SI{7.3}{\percent}
(cf.~\Cref{fig:hh_branching_ratios}) and the distinct signature of
\tauleptons. Previous searches in the \bbtautau channel were performed by the
ATLAS and CMS collaborations using data recorded in
Run~1~\cite{HIGG-2013-33,CMS-HIG-15-013} and at the beginning of Run~2 of the
LHC~\cite{HIGG-2016-16-witherratum,CMS-HIG-17-002}.  Other searches for Higgs
boson pair production using the full Run~2 \pp~collision datasets were performed
by the ATLAS and CMS collaborations in the \bbtautau~\cite{CMS-HIG-20-010},
\bbbb~\cite{HDBS-2019-29,CMS-HIG-20-005}, and
\bbyy~\cite{HDBS-2018-34,CMS-HIG-19-018} channels.
% The context for the results presented in this chapter is provided by other
% searches for Higgs boson pair production by the ATLAS and CMS collaborations
% using \pp~collision datasets recorded in Run~2 of the LHC in the
The work presented in this chapter contributes to
Refs.~\cite{ATLAS-CONF-2021-030,HDBS-2018-40}.

The search for SM~\HH production aims to set upper limits on the signal strength
and cross section of SM~\HH production via \ggF and VBF provided no significant
deviation from the background-only hypothesis is observed. Under the SM
expectation, the size of the \pp~collision dataset recorded during Run~2 of the
LHC is not sufficient to obtain evidence for SM~\HH production.  Nevertheless,
BSM physics could enhance the non-resonant production of Higgs boson pairs,
which can be probed in this search.
% with kinematic properties similar to those of the SM~\HH signal being probed
% in this search.
Interpretations of the SM~\HH search in terms of anomalous values of the Higgs
boson self-coupling strength are not subject of this chapter but rather of
\Cref{sec:higgs_self_coupling}. The search for resonant \HH production targets
scalar resonances with masses ranging from \SIrange{251}{1600}{\GeV} produced
via \ggF and aims to, in the absence of significant deviations from the SM
expectation, set upper limits on the production cross section of the scalar
resonance.

This chapter is structured as follows: \Cref{sec:data_and_simulation} describes
the data and simulated event samples used for this search. The reconstruction of
physics objects and selection of events is discussed in
\Cref{sec:object_reconstruction,sec:event_selection}, respectively. The
estimation of background processes contributing to the search is described in
\Cref{sec:background_estimation} with a focus on the estimation of backgrounds
with quark- or gluon-initiated jets that are misidentified as \tauhadvis. The
multivariate analysis used to distinguish signal from background events is
discussed in \Cref{sec:multivariate_analysis}. Systematic uncertainties and the
statistical interpretation are described in
\Cref{sec:uncertainties,sec:statistical_analysis}. This chapter concludes with a
discussion of the results in \Cref{sec:result_discussion} including a comparison
with other searches for Higgs boson pair production performed by the ATLAS and
CMS collaborations as well as an outlook on future prospects of SM~\HH searches.
% at the High-Luminosity LHC (HL-LHC).


%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../../phd_thesis"
%%% End:
