The recommendations for the derivation of uncertainties on the modelling of
selected physics processes by the ATLAS collaboration are summarised in the
following. Modelling uncertainties for \ttbar and single top-quark processes are
summarised in \Cref{app:top_uncertainties}. Uncertainties on the modelling of
\Zjets processes are summarised in \Cref{app:zjets_uncertainties}.

\todo[inline]{Make this more generic.}


\subsection{Uncertainties on the Modelling of \ttbar and Single Top Production
  in Simulation}%
\label{app:top_uncertainties}

% https://twiki.cern.ch/twiki/bin/viewauth/AtlasProtected/PmgTopProcesses
Theoretical uncertainties on the acceptance of \ttbar in simulation, which uses
\POWHEGBOX[v2]~\cite{Frixione:2007nw} as a matrix element (ME) generator
interfaced to \PYTHIA[8.230]~\cite{Sjostrand:2014zea} for the parton shower
(PS), need to be estimated. The cross section of \ttbar is allowed to vary
freely in the fit model, therefore uncertainties only affecting the
normalisation of \ttbar are omitted. However, uncertainties changing the
relative acceptance between regions (i.e.\ uncertainties depending on \tauhadvis
\pT or decay mode) entering the fit and effects changing the shape of the \mTW
discriminant need to be estimated. The following sources of uncertainties are
considered:
\begin{description}

\item[Hard scatter simulation and NLO+PS matching] The uncertainty on the
  modelling due to the choice of generator describing the hard scatter process
  is estimated by comparing the nominal simulation of \ttbar with an alternative
  setup replacing \POWHEGBOX[v2] with \MGNLO as the ME generator. This
  comparison serves to probe the effect of different matching schemes between
  the NLO matrix element and parton shower employed by both
  generators.\footnote{\POWHEGBOX[v2] uses the \POWHEG
    method~\cite{Nason:2004rx,Frixione:2007vw,Alioli:2010xd} and \MGNLO the
    MC@NLO method~\cite{Frixione:2002ik} for NLO+PS matching.}

\item[Parton shower and hadronisation model] The uncertainty of the modelling of
  the parton shower and non-perturbative effects is estimated by replacing
  \PYTHIA[8] with \HERWIG[7.0.4] as the PS generator.

\item[Missing higher order contributions] The renormalisation scale~(\muR) and
  factorisation scale~(\muF) is doubled (halved) to probe the effect of
  truncating the perturbative expansion in \alphas when simulating the hard
  scatter process. Perturbative QCD calculations to sufficiently high order
  should be approximately independent of the choice of scale.

\item[Initial and final state radiation (ISR / FSR)] An uncertainty on the
  emission of ISR is estimated by varying $\alphas^\text{ISR}$ in the A14 set of
  tuned parameters for \PYTHIA[8]~\cite{ATL-PHYS-PUB-2014-021}.
  % Tune of the MPI, ISR, FSR parameters in Pythia8
  An estimate of the uncertainty from the modelling of FSR emissions is
  estimated by doubling (halving) the renormalisation scale used for FSR
  branchings in
  \PYTHIA[8]~\cite{Sjostrand:2014zea,Mrenna:2016sih,pythia-variations-online}.
  % https://pythia.org/latest-manual/Variations.html

\item[Damping factor for additional emissions] The damping parameter~\hdamp in
  \POWHEGBOX[v2] is increased to $3 m_\text{top}$ (from $1.5 m_\text{top}$) and
  compared with the nominal simulation setup. The \hdamp parameter controls the
  transverse momentum of additional radiation when matching \POWHEGBOX[v2] to
  \PYTHIA[8] using the
  \POWHEG-method~\cite{ATL-PHYS-PUB-2016-020,ATL-PHYS-PUB-2020-023}.

\end{description}
% PDF uncertainties were neglected.
The prescriptions to calculate the \ttbar modelling uncertainties are a revised
version of the methodology outlined in Ref.~\cite{ATL-PHYS-PUB-2020-023} for
analyses targeting the \pp~collision dataset collected with the ATLAS detector
during Run~2 of the LHC.

\todo[inline]{PDF + \alphas ?}

Variations of the renormalisation and factorisation scales used for the ME
generation are provided by internal re-weighting in \POWHEGBOX[v2]. Similarly,
\PYTHIA[8] provides variations of initial and final state emissions by varying
the renormalisation scales in the PS by
re-weighting~\cite{Mrenna:2016sih,pythia-variations-online}. This approach
allows to estimate uncertainties without changing the particle-level predictions
of the simulation program, thus avoiding the need to re-run the detector
simulation.

\todo[inline]{Note option of combining the initial state radiation and scale
  variations uncertainties in old recommendations.}


\subsection{Uncertainties on the Modelling of \Zjets production in Simulation}%
\label{app:zjets_uncertainties}

The relative acceptance uncertainties originating from the modelling of the \ZHF
background in simulation are estimated by performing variations of the event
generation process. A brief description of the variations is given in the
following, reproducing the prescriptions developed by the ATLAS
collaboration:\todo{Move to appendix}
\begin{description}

\item[Factorisation and renormalisation scales] Six variations of the
  factorisation and renormalisation scales are performed using internal
  re-weighting implemented in \SHERPA[2.2.1]~\cite{Bothmann:2019yzt}, altering
  the scales by factors of $\frac{1}{2}$ and $2$. The following variations are
  considered:
  \begin{align*}
    \left( \frac{\muF}{\muF^{\text{nom.}}}, \frac{\muR}{\muR^{\text{nom.}}} \right) \in
    \left\{ (\tfrac{1}{2}, \tfrac{1}{2}), (\tfrac{1}{2}, 1), (1, \tfrac{1}{2}), (1, 2), (2, 1), (2, 2) \right\} \,\text{,}
  \end{align*}
  where $\muF^{\text{nom.}}$ and $\muR^{\text{nom.}}$ are the nominal values of
  the scales.

\item[Resummation scale] The scale of the resummation of soft gluon emissions in
  the \SHERPA parton shower is varied by factors of $\frac{1}{2}$ and
  2. Variations of the resummation scale are provided in parameterised form with
  respect to the default \SHERPA configuration in Ref.~\cite{anders:2017}.

\item[Multi-jet merging scale] The simulation of \Zjets events with
  \SHERPA[2.2.1] uses matrix elements of NLO accuracy for up to two and LO for
  up to four partons. These multi-parton matrix elements are merged with the
  parton shower using an extension of the CKKW
  algorithm~\cite{Catani:2001cc,Hoeche:2009rj,Hoeche:2012yf}. The characteristic
  scale~$Q_{\text{cut}}$ of the multi-jet merging algorithm is varied from its
  nominal value of $Q_{\text{cut}} = \SI{20}{\GeV}$ to \SI{15}{\GeV} and
  \SI{30}{\GeV}~\cite{anders:2017}. These variations are provided, following the
  approach for the resummation scale, in parameterised form in
  Ref.~\cite{anders:2017}.
  %\cite{ATLAS:2021yza}

\item[PDF+\alphas and PDF choice] Uncertainties on the \NNPDF[3.0nnlo] set of
  PDFs~\cite{Ball:2014uwa} are evaluated using 100 replica sets provided through
  the \textsc{LHAPDF6} library~\cite{Buckley:2014ana} and implemented using
  internal re-weighting in \SHERPA. The uncertainty on \alphas propagated by
  comparing \NNPDF[3.0nnlo] PDF sets with $\alphas(\mZ^2) = 0.117$ and $0.119$
  with the nominal set using a value of $0.118$. Finally, an uncertainty on the
  choice of PDF set is estimated by comparing with two alternative PDF sets
  \MMHT[nnlo68cl]~\cite{Harland-Lang:2014zoa} and
  \CT[14nnlo]~\cite{Dulat:2015mca}.

\item[Alternative generator and parton shower] The prediction of \Zjets with the
  default configuration of \SHERPA[2.2.1] is compared to an alternative setup
  using~\MGNLO[2.2.2]~\cite{Alwall:2014hca} for the calculation of the hard
  interaction at LO interfaced to~\PYTHIA[8.186]~\cite{Sjostrand:2007gs} for
  parton showering.

\end{description}
Relative acceptance uncertainties on the \ZHF background are derived using these
variations according to~\Cref{eq:relative_acceptance_uncertainty} for the three
signal regions. The uncertainties are summarised
in~\Cref{tab:uncertainties_zhf_extrapol} separately for all sources of
uncertainty.


%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../../phd_thesis"
%%% End:
